Contextual Identity: Freedom to be All Your Selves

### 2013-04-24 paper.tex
minor cleanups

The desire for privacy balances the desire for spontaneous, positive human
interaction, which often requires sharing personal information. The act of
sharing alone does not negate expectation of privacy~\cite{boyd2}.
%The desire for spontaneous, positive human interaction often requires sharing
%personal information. Sharing information can lead to loss of privacy, though
%it does not negate expectation of privacy~\cite{boyd2}. 
Helen Nissenbaum has long argued that privacy violations come not from the
simple sharing of personal information, but rather from sharing that
information in a way that violates social norms~\cite{nissenbaum}.

Lesbian, gay, bisexual, and transgendered youth are at risk for contextual
identity violations.  Access to support is crucial for LGTB-identified youth,
who are at higher risk for bullying and suicide~\cite{hrc}, and
much of that support is only accessible through the Internet in socially
conservative areas. In 2011 Bobbi Duncan and Taylor McCormick, two
students at the University of Texas at Austin, were inadvertently outed when
their choir director added them to a Facebook group for Queer
Chorus~\cite{fowler}. Even though both students had blocked their parents from
seeing any personal posts revealing their sexuality, Queer Chorus is a
public group and so Facebook published notifications to Duncan's and McCormick's
friends, including their parents. Duncan later attempted suicide~\cite{duncan}.

@@

Barth et al.~provide a formal model for contextual
integrity of user data~\cite{barth}. Their work models
the flow of personal information in terms of knowledge states held by agents
acting in given roles, and works well for expressing privacy legislation such as
HIPPA and COPPA. Two limitations of this work are that it does not
model information where the subject involves more than one person, and it
only takes the type of information into account rather than the content or tone
of the information. For example, ``Bob told Alice that Charlie cooked a
delicious dinner'' is out of scope of the model.  In the context of social
networks, information is rarely about a single person, and the tone of the
message is as important to user effect as the contents.

@@

Evidence suggests that a person's major privacy concerns come from others he or she
knows: friends, family, and co-workers~\cite{fbtips2,fbtips1}.  Surveys
conducted by Wang et al.~suggest that typical regrets from posting on Facebook
stem almost {\em exclusively} from fear of negative interactions with people the individual knows~\cite{wang}. The consequences of these negative interactions can
lead to loss of employment or breaking personal relationships.  Most previous
privacy research has focused on distant adversaries, where the bad actor is a
behavioral tracking service, state agent or unknown eavesdropper; in contrast,
we assert that people care more about their interactions with others close to
them.

@@

Although access control and usability play a role in contextual identity
management, they are not a panacea: many of the below examples fall out of
scope of traditional access control models.

\label{sec:examples}
\subsection{Redistributing information out of context}
Redistributing information out of its original context often leads to
embarrassment~\cite{nissenbaum}.  In November 2007, Facebook Beacon allowed
third-party sites to publish purchases, travel bookings, movie rentals and more
to the user's activity stream.  Difficulty of opt-out and lack of visibility
into what was being published led to user surprise and dismay~\cite{mccarthy,nakashima}.  In December 2007, Google Reader exposed
RSS feeds of user-marked news stories to the user's Google Talk contacts, which
includes everyone with whom a user has chatted whether they be co-workers,
supervisors, or friends. Although this feed was always public, prior to this
launch it was not discoverable, leading to a bad experience for many
users~\cite{helft}. In February 2010, Google Buzz launched, a product that
exposed the user's most frequent Gmail and Google Talk contacts and their
publicly available (through previously less discoverable) news and photos.  The
combination of exposing contacts (which could include abusive ex-husbands,
co-workers, and friends), as well as aggressively linking photos and news
items, resulted in a huge backlash~\cite{fugitivus,buzz}. In September 2012,
Facebook imported old wall posts from 2008 into the new Timeline interface.
Although wall posts were always visible from profile pages, the new Timeline
interface brought old wall posts, which people used to treat as private messages
before the advent of ``Like'' and comment buttons, to the attention of an
audience that the posting party never anticipated~\cite{timeline}.

\subsection{Policy errors}
\label{sec:policies}
Some service providers have policies which preclude isolating multiple
contextual identities. For example, Facebook and Google have a ``Real Names''
policy, which requires people to register for accounts with their legal
name.\footnote{Facebook's policy: \url{http://www.facebook.com/help/?page=258984010787183}, Google's policy: \url{http://support.google.com/plus/bin/answer.py?hl=en&answer=1228271}}
%\cite{fb_names,google_names}.  
These policies ignore that
community-building happens in many different contexts, that individuals have
legitimate reasons for presenting different identities in different contexts,
and they don't necessarily want those identities to be linked. For
example, disallowing avatar handles as a primary identifier makes building a
gaming community difficult. % Linking a gaming identity to a professional
%identity has already caused problems for 2012 Maine Senate candidate and World
%of Warcraft gamer Colleen Lachowicz~\cite{maine}.
It is impossible to isolate multiple contextual identities in these networks
without violating the terms of service.

@@

\subsection{User interface errors}
It is all too easy for users to broadcast information to
an unintended audience. This phenomenon is so common on Twitter that it has its
own name, ``DM fail'', or Direct Message fail: when the user posts a public
message instead of a private, direct message. Representative Anthony Weiner was
a victim of this mistake when he inadvertently published compromising pictures
of himself~\cite{weiner}. Considering that this mistake requires
mistyping a single character (\texttt{@} instead of \texttt{d}), it's no
surprise that DM failures are so common.

>>>>

The desire for privacy balances the desire for spontaneous, positive human
interaction and sharing personal information. The act of
sharing alone does not negate expectation of privacy~\cite{boyd2}.
%The desire for spontaneous, positive human interaction often requires sharing
%personal information. Sharing information can lead to loss of privacy, though
%it does not negate expectation of privacy~\cite{boyd2}. 
Helen Nissenbaum has long argued that privacy violations come not from the
simple sharing of personal information itself, but rather from sharing that
information in a way that violates social norms~\cite{nissenbaum}.

For example, lesbian, gay, bisexual, and transgendered youth are particularly at risk for contextual
identity violations.  Access to support is crucial for LGTB-identified youth,
who are at higher risk for bullying and suicide~\cite{hrc}, and
much of that support is only accessible through the Internet in socially
conservative areas. In 2011 Bobbi Duncan and Taylor McCormick, two
students at the University of Texas at Austin, were inadvertently outed when
their choir director added them to a Facebook group for Queer
Chorus~\cite{fowler}. Even though both students had blocked their parents from
seeing any personal posts revealing their sexuality, Queer Chorus is a
public group and so Facebook published notifications to Duncan's and McCormick's
friends, including their parents. Duncan later attempted suicide~\cite{duncan}.

@@

Barth et al.~provide a formal model for contextual
integrity of user data~\cite{barth}. Their work models
the flow of personal information in terms of knowledge states held by agents
acting in given roles, and works well for expressing privacy legislation such as
HIPPA and COPPA. Two limitations of this work are that it does not
model information where the subject involves more than one person, and it
only takes the type of information into account rather than the content or tone
of the information. For example, ``Bob told Alice that Charlie cooked a
delicious dinner'' is out of scope of the model.  In the context of social
networks, information is rarely about a single person, and the tone of the
message is as important to its effect on a user as the contents.

@@

Evidence suggests that a person's major privacy concerns come from others he or she
knows: friends, family, and co-workers~\cite{fbtips2,fbtips1}.  Surveys
conducted by Wang et al.~suggest that typical regrets from posting on Facebook
stem almost {\em exclusively} from fear of negative interactions with people the individual knows~\cite{wang}. The consequences of these negative interactions can
lead to loss of employment or breaking personal relationships.  Most previous
privacy research has focused on distant adversaries, where the bad actor is a
behavioral tracking service, state agent or unknown eavesdropper; in contrast,
we assert that people are concerned about their interactions with others close to
them.

@@

Although additional access control and improved usability can improve contextual identity
management, they are not a panacea: many of the below examples fall out of
scope of traditional access control models.

\label{sec:examples}
\subsection{Redistributing information in a different context}
Redistributing information out of its original context often leads to
embarrassment~\cite{nissenbaum}.  In November 2007, Facebook Beacon allowed
third-party sites to publish purchases, travel bookings, movie rentals and more
to the user's activity stream.  Difficulty of opt-out and lack of visibility
into what was being published led to user surprise and dismay~\cite{mccarthy,nakashima}.  In December 2007, Google Reader exposed
RSS feeds of user-marked news stories to the user's Google Talk contacts, which
includes everyone with whom a user has chatted whether they be co-workers,
supervisors, or friends. Although this feed was always public, prior to this
launch it was not discoverable, leading to a bad experience for many
users~\cite{helft}. In February 2010, Google Buzz launched, a product that
exposed the user's most frequent Gmail and Google Talk contacts and their
publicly available (through previously less discoverable) news and photos.  The
combination of exposing contacts (which could include abusive ex-husbands,
co-workers, and friends), as well as aggressively linking photos and news
items, resulted in a huge backlash~\cite{fugitivus,buzz}. In September 2012,
Facebook imported old wall posts from 2008 into the new Timeline interface.
Although wall posts were always visible from profile pages, the new Timeline
interface brought old wall posts, which people used to treat as private messages
before the advent of ``Like'' and comment buttons, to the attention of an
audience that the posting party never anticipated~\cite{timeline}.

\subsection{Unaccommodating policies}
\label{sec:policies}
Some service providers have policies which preclude isolating multiple
contextual identities. For example, Facebook and Google have a ``Real Names''
policy, which requires people to register for accounts with their legal
name.\footnote{Facebook's policy: \url{http://www.facebook.com/help/?page=258984010787183}, Google's policy: \url{http://support.google.com/plus/bin/answer.py?hl=en&answer=1228271}}
%\cite{fb_names,google_names}.  
These policies ignore that
community-building happens in many different contexts, that individuals have
legitimate reasons for presenting different identities in different contexts,
and they don't necessarily want those identities to be linked. For
example, disallowing avatar handles as a primary identifier makes building a
gaming community difficult. % Linking a gaming identity to a professional
%identity has already caused problems for 2012 Maine Senate candidate and World
%of Warcraft gamer Colleen Lachowicz~\cite{maine}.
It is impossible to isolate multiple contextual identities in these networks
without violating the terms of service.

@@

\subsection{Confusing user interfaces}
It is all too easy for users to broadcast information to
an unintended audience. This phenomenon is so common on Twitter that it has its
own name, ``DM fail'', or Direct Message fail: when the user posts a public
message instead of a private, direct message. Representative Anthony Weiner was
a victim of this mistake when he inadvertently published compromising pictures
of himself~\cite{weiner}. Considering that this mistake requires
mistyping a single character (\texttt{@} instead of \texttt{d}), it's no
surprise that DM failures are so common.
